---
title: "HM_figures"
author: "Abby Lewis"
date: "2026-01-20"
output: html_document
---

TO DO: 
- Interpolate fluxes
- Bring in EVI

Hot moments are important across all chambers (Figure 2)
The importance of hot moment events increases with median flux (Figure 3)
Because of hot moments, manual sampling underestimates fluxes (Figure 4)
Hot moments occur as water level rises (Figure 2, maybe also another)
However, the magnitude of methane emitted is poorly predicted by the magnitude of change in water level (Figure 5)

Load data and packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggridges)
source("../analysis/hm_style.R")

part <- read_csv(here::here("processed_data", "L2- partitioned_and_gap_filled.csv")) %>%
  rename(DateTime = TIMESTAMP) %>%
  mutate(CH4_umol_m2_h = CH4_filled * (60 * 60),
         MIU_VALVE = factor(MIU_VALVE, levels = 1:12, 
                            labels = chamber_levels2),
         date = as.Date(DateTime)) %>%
  filter(CH4_filled < 1) #Removing n = 1 flux from chamber 12. Probably good to add an si figure showing what happens if we don't do this

all <- read_csv("../processed_data/L0_for_dashboard.csv")
met <- read_csv("../processed_data/met_2025_dashboard.csv") %>%
  mutate(Salinity = ifelse(TIMESTAMP > as_datetime("2025-07-22 12:00:00") &
                             TIMESTAMP < as_datetime("2025-07-24 12:00:00") , 
                           NA, Salinity),
         Salinity = ifelse(Salinity < 1 , 
                           NA, Salinity))
wl <- read_csv("../processed_data/water_level_dashboard.csv")
evi <- read_csv("../processed_data/evi.csv")

```

Check qaqc

```{r}
p <- all %>%
  filter(date >= Sys.Date() - days(700)) %>%
  mutate(MIU_VALVE = factor(MIU_VALVE, levels = 1:12, 
                            labels = chamber_levels2)) %>%
  ggplot(aes(x = flux_start, y = CH4_init, color = MIU_VALVE))+
  geom_line()

plotly::ggplotly(p)

all %>%
  filter(date >= Sys.Date() - days(700)) %>%
  mutate(MIU_VALVE = factor(MIU_VALVE, levels = 1:12, 
                            labels = chamber_levels2)) %>%
  ggplot(aes(x = CH4_init, y = CH4_slope_ppm_per_day, color = MIU_VALVE))+
  geom_point(size = 0.5)

all %>%
  arrange(flux_start) %>%
  mutate(init_change = CH4_init - lag(CH4_init)) %>%
  filter(date >= Sys.Date() - days(700)) %>%
  mutate(MIU_VALVE = factor(MIU_VALVE, levels = 1:12, 
                            labels = chamber_levels2)) %>%
  ggplot(aes(x = init_change, y = CH4_slope_ppm_per_day, color = MIU_VALVE))+
  geom_point(size = 0.5)

```


Data summary (Figure 2)

```{r}
min_annual = min(part$date[!is.na(part$CH4_umol_m2_h)])
max_annual = max(part$date[!is.na(part$CH4_umol_m2_h)])

annual <- part %>%
  group_by(date, MIU_VALVE) %>%
  summarize(CH4_umol_m2_h = mean(CH4_umol_m2_h, na.rm = T)) %>%
  ggplot(aes(x = date, y = CH4_umol_m2_h, color = MIU_VALVE))+
  geom_hline(yintercept = 0, color = "grey80")+
  geom_line(size = 0.5)+
  coord_cartesian(xlim = c(as.Date(min_annual), as.Date(max_annual)))+
  scale_color_manual(values = color.gradient, name = "Chamber")+
  theme_hm +
  theme(axis.title.x = element_blank())+
  ylab(expression(atop(CH[4]~"flux","(µmol"~m^-2~h^-1*")")))

annual_AT <- met %>%
  mutate(date = as.Date(TIMESTAMP)) %>%
  group_by(date) %>%
  summarize(AirTC_Avg = mean(AirTC_Avg, na.rm = T)) %>%
  ggplot(aes(x = date, y = AirTC_Avg))+
  geom_hline(yintercept = 0, color = "grey80")+
  geom_line(size = 0.5, color = "grey20")+
  coord_cartesian(xlim = c(as.Date(min_annual), as.Date(max_annual)))+
  scale_color_manual(values = color.gradient, name = "Chamber")+
  theme_hm +
  theme(axis.title.x = element_blank())+
  ylab(expression(atop("Air temp."," (ºC)")))

annual_wl <- wl %>%
  mutate(date = as.Date(TIMESTAMP)) %>%
  group_by(date) %>%
  summarize(Depth_cm = mean(Depth_cm, na.rm = T)) %>%
  ggplot(aes(x = date, y = Depth_cm))+
  geom_hline(yintercept = 0, color = "grey80")+
  geom_line(size = 0.5, color = "grey20")+
  coord_cartesian(xlim = c(as.Date(min_annual), as.Date(max_annual)))+
  scale_color_manual(values = color.gradient, name = "Chamber")+
  theme_hm +
  theme(axis.title.x = element_blank())+
  ylab(expression(atop("Water level"," (cm)")))

gpp <- part %>%
  group_by(date, MIU_VALVE) %>%
  summarize(GPP = sum(GPP, na.rm = T)) %>%
  ggplot(aes(x = date, y = GPP, color = MIU_VALVE))+
  geom_hline(yintercept = 0, color = "grey80")+
  geom_line(size = 0.5)+
  coord_cartesian(xlim = c(as.Date(min_annual), as.Date(max_annual)))+
  scale_color_manual(values = color.gradient, name = "Chamber")+
  theme_hm +
  theme(axis.title.x = element_blank())+
  ylab(expression(atop("Daily mean"~CH[4]~"flux","(µmol"~m^-2~h^-1*")")))

resp <- part %>%
  group_by(date, MIU_VALVE) %>%
  summarize(R = sum(Reco, na.rm = T)) %>%
  ggplot(aes(x = date, y = R, color = MIU_VALVE))+
  geom_hline(yintercept = 0, color = "grey80")+
  geom_line(size = 0.5)+
  coord_cartesian(xlim = c(as.Date(min_annual), as.Date(max_annual)))+
  scale_color_manual(values = color.gradient, name = "Chamber")+
  theme_hm +
  theme(axis.title.x = element_blank())+
  ylab(expression(atop("Daily mean"~CH[4]~"flux","(µmol"~m^-2~h^-1*")")))

annual_evi <- evi %>%
  ggplot(aes(x = Date, y = evi_predicted))+
  geom_line(size = 0.5)+
  geom_point(aes(y = evi_observed))+
  theme_hm +
  coord_cartesian(xlim = c(as.Date(min_annual), as.Date(max_annual)))+
  theme(axis.title.x = element_blank())+
  ylab(expression(atop("EVI","")))

start_date <- "2025-03-27"
end_date <- "2025-04-06"

focal_ch4 <- part %>%
  filter(date %in% seq.Date(as.Date(start_date), as.Date(end_date), by = "1 day")) %>%
  ggplot(aes(x = DateTime, y = CH4_umol_m2_h, color = MIU_VALVE))+
  geom_hline(yintercept = 0, color = "grey80")+
  geom_line(size = 0.5)+
  scale_color_manual(values = color.gradient, name = "Chamber")+
  theme_hm +
  theme(axis.title.x = element_blank())+
  ylab(expression(atop(CH[4]~"flux ", "(µmol"~m^-2~h^-1*")")))

focal_wl <- wl %>%
  mutate(date = as.Date(TIMESTAMP)) %>%
  filter(date %in% seq.Date(as.Date(start_date), as.Date(end_date), by = "1 day")) %>%
  ggplot(aes(x = TIMESTAMP, y = Depth_cm))+
  geom_hline(yintercept = 0, color = "grey80")+
  geom_line(size = 0.5, color = "grey20")+
  scale_color_manual(values = color.gradient, name = "Chamber")+
  theme_hm +
  theme(axis.title.x = element_blank())+
  ylab(expression(atop("Water level"," (cm)")))

focal_AT <- met %>%
  mutate(date = as.Date(TIMESTAMP)) %>%
  filter(date %in% seq.Date(as.Date(start_date), as.Date(end_date), by = "1 day")) %>%
  ggplot(aes(x = TIMESTAMP, y = AirTC_Avg))+
  geom_hline(yintercept = 0, color = "grey80")+
  geom_line(size = 0.5, color = "grey20")+
  scale_color_manual(values = color.gradient, name = "Chamber")+
  theme_hm +
  theme(axis.title.x = element_blank())+
  ylab(expression(atop("Air temp."," (ºC)")))

comb <- ggpubr::ggarrange(
  annual+
    guides(color = guide_legend(ncol = 4))+
    theme(legend.position = "bottom",
          legend.title.position = "top"), 
  patchwork::wrap_plots(focal_ch4 +
                      scale_x_datetime(date_breaks = "5 day",
                                       date_labels = "%d %b")+
                        theme(axis.text.x = element_blank(),
                              axis.ticks.x = element_blank(),
                              legend.position = "none"), 
                    focal_wl +
                      scale_x_datetime(date_breaks = "5 day",
                                       date_labels = "%d %b")+
                        theme(axis.text.x = element_blank(),
                              axis.ticks.x = element_blank(),
                              legend.position = "none"), 
                    focal_AT +
                      scale_x_datetime(date_breaks = "5 day",
                                       date_labels = "%d %b")+
                        theme(legend.position = "none"), 
                    ncol = 1), 
  ncol = 2, widths = c(1.7,1), labels = c("a", "b"),
  legend = "bottom", common.legend = T
)
comb
ggsave("../figures/Figure 2 - Data.jpg", dpi = 300, width = 6, height = 4)

comb <- patchwork::wrap_plots(
  a = patchwork::wrap_plots(
    annual +
      guides(color = guide_legend(ncol = 4)) +
      theme(axis.text.x = element_blank(),
            axis.ticks.x = element_blank(),
            legend.position = "none") +
      ggtitle("Seasonal (daily average)"), 
    annual_wl +
      theme(axis.text.x = element_blank(),
            axis.ticks.x = element_blank(),
            legend.position = "none"), 
    annual_AT, 
    ncol = 1, design = c("A
                          A
                          B
                          C")
  ),
  b = patchwork::wrap_plots(
    focal_ch4 +
      scale_x_datetime(date_breaks = "5 day", date_labels = "%d %b") +
      theme(axis.text.x = element_blank(),
            axis.ticks.x = element_blank(),
            axis.title.y = element_blank(),
            legend.position = "none") +
      ggtitle("Short-term (hourly)"), 
    focal_wl +
      scale_x_datetime(date_breaks = "5 day", date_labels = "%d %b") +
      theme(axis.text.x = element_blank(),
            axis.ticks.x = element_blank(),
            axis.title.y = element_blank(),
            legend.position = "none"), 
    focal_AT +
      scale_x_datetime(date_breaks = "5 day", date_labels = "%d %b") +
      theme(legend.position = "none",
            axis.title.y = element_blank()),
    ncol = 1,
    design = c("A
               A
               B
               C")
  ),
  ncol = 2, widths = c(2,1),
  guides = "collect"
) + 
  patchwork::plot_layout(guides = "collect") +
  theme(legend.position = "bottom", legend.justification = "center", 
        legend.title.position = "top")

comb

ggsave("../figures/Figure 2 - Data drivers.jpg", dpi = 300, width = 7, height = 5)

#Why do our fluxes go up as WL goes up
part %>%
  filter(MIU_VALVE == "Ch. 11 (+5.25 ºC)",
         month(DateTime) == 10,
         day(DateTime) %in% 10:14)%>%
  select(DateTime, Depth_cm, Salinity, CH4_filled) %>%
  pivot_longer(c(Depth_cm, Salinity, CH4_filled)) %>%
  ggplot(aes(x = DateTime, y = value))+
  geom_line()+
  facet_wrap(~name, scales = "free_y", ncol = 1)

part %>%
  filter(MIU_VALVE == "Ch. 11 (+5.25 ºC)",
         month(DateTime) == 7,
         day(DateTime) %in% 22:30)%>%
  select(DateTime, Depth_cm, Salinity, CH4_filled) %>%
  pivot_longer(c(Depth_cm, Salinity, CH4_filled)) %>%
  ggplot(aes(x = DateTime, y = value))+
  geom_line()+
  facet_wrap(~name, scales = "free_y", ncol = 1)

part %>%
  filter(MIU_VALVE == "Ch. 11 (+5.25 ºC)")%>%
  select(DateTime, Depth_cm, Salinity, CH4_filled) %>%
  pivot_longer(c(Depth_cm, Salinity, CH4_filled)) %>%
  ggplot(aes(x = DateTime, y = value))+
  geom_line()+
  facet_wrap(~name, scales = "free_y", ncol = 1)
```

Quants

```{r}
quants2_data <- part %>%
  filter(!is.na(CH4_filled)) %>%
  select(-CH4) %>%
  rename(CH4 = CH4_filled) %>%
  mutate(Date = as.Date(DateTime),
         Temp_C = sub(" ºC", "", str_extract(as.character(MIU_VALVE), "[0-9]+\\.*[0-9]* ºC")),
         Temp_C = as.numeric(sub("mb", "0", Temp_C))) %>%
  group_by(Date)  %>%
  filter(length(unique(MIU_VALVE)) == 12) %>%
  group_by(MIU_VALVE, Temp_C) %>%
  summarize(q_50 = quantile(CH4, .50),
            #q_5 = quantile(CH4, .05),
            q_95 = quantile(CH4, .95),
            q_99 = quantile(CH4, .99)) %>%
  pivot_longer(q_50:q_99) %>%
  group_by(name) %>%
  mutate(ref = mean(value[Temp_C == 0]),
         name = as.numeric(sub("q_", "", name)),
         name_txt = paste0(name, "th quantile"),
         name_txt = fct_reorder(name_txt, name)) 

quants_plot <- quants2_data %>%
  mutate(pct = value/ref*100,
         dif = value-ref) %>%
  ungroup() %>%
  select(-value, -ref, -name) %>%
  pivot_longer(c(pct, dif)) %>%
  mutate(name = factor(name, 
                       levels = c("dif", "pct"),
                       labels = c("'Difference (µmol'~m^-2~h^-1*`)`", "Percent")),
         name_txt = paste0("'", name_txt, "'")) %>%
  ggplot(aes(x = Temp_C, y = value))+
  geom_hline(yintercept = 0, color = "grey50")+
  geom_smooth(color = "black", fill = "grey70", method = "lm")+
  geom_point(aes(fill = MIU_VALVE), color = "black", 
             shape = 21, size = 2, stroke = 0.5)+
  scale_fill_manual(values = color.gradient, name = "Chamber")+
  ylab(expression(CH[4]~"flux relative to ambient"))+
  xlab("Temperature treatment (ºC above ambient)")+
  guides(fill = guide_legend(ncol = 3))+
  facet_grid(name~name_txt, scales = "free", labeller = "label_parsed", 
             switch = "y")+
  theme_hm +
  theme(strip.placement = "outside")

quants_plot

ggsave(quants_plot, filename = "../figures/Quants.jpg", dpi = 300, width = 4, height = 5)
```


The importance of hot moment events increases with median flux (Figure 3)

```{r}
# Tried using R pacakage, but I'm skeptical of this because only a normal 
# distribution is implemented currently.
# devtools::install_github("https://github.com/jonathan-walter/hotspomoments")
#library(hotspomoments)
#
#hshmtest(part$CH4_umol_m2_h[!is.na(part$CH4_umol_m2_h)], "skewness")
#hshm_id <- hshmid(part$CH4_umol_m2_h[!is.na(part$CH4_umol_m2_h)], 
#                  criteria = "ref.normal", side="upper", thresh=0.95)
##Plot where in the distribution the HSHM are identified
#hist(part$CH4_umol_m2_h[!is.na(part$CH4_umol_m2_h)])
#points(part$CH4_umol_m2_h[!is.na(part$CH4_umol_m2_h)][hshm_id], 
#       rep(0,sum(hshm_id)), pch=19, col="red")
#
##Inspect HSHM observations
#hshm <- part[!is.na(part$CH4_umol_m2_h),][hshm_id,]

hm_df <- part %>%
  filter(!is.na(CH4_umol_m2_h)) %>%
  group_by(date) %>%
  filter(length(unique(MIU_VALVE)) == 12)

hot_moments <- hm_df %>%
  group_by(MIU_VALVE) %>%
  mutate(hot_moment_quant_998 = ifelse(CH4_umol_m2_h >= 
                                     quantile(CH4_umol_m2_h, 
                                              probs = 0.998, na.rm = TRUE),
                                   "hot", "not"),
         hot_moment_quant_99 = ifelse(CH4_umol_m2_h >= 
                                     quantile(CH4_umol_m2_h, 
                                              probs = 0.99, na.rm = TRUE),
                                   "hot", "not"),
         hot_moment_quant_95 = ifelse(CH4_umol_m2_h >= 
                                     quantile(CH4_umol_m2_h, 
                                              probs = 0.95, na.rm = TRUE),
                                   "hot", "not"),
         hot_moment_iqr = ifelse(CH4_umol_m2_h > 
                                     (1.5 * (quantile(CH4_umol_m2_h, 
                                              probs = 0.75, na.rm = TRUE) - 
                                        quantile(CH4_umol_m2_h, 
                                              probs = 0.25, na.rm = TRUE))),
                                   "hot", "not")) %>%
  summarize(median = median(CH4_umol_m2_h, na.rm = T),
            mean = mean(CH4_umol_m2_h, na.rm = T),
            sum = sum(CH4_umol_m2_h, na.rm = T),
            hot_sum_998 = sum(CH4_umol_m2_h[hot_moment_quant_998 == "hot"], na.rm = T),
            hot_sum_99 = sum(CH4_umol_m2_h[hot_moment_quant_99 == "hot"], na.rm = T),
            hot_sum_95 = sum(CH4_umol_m2_h[hot_moment_quant_95 == "hot"], na.rm = T),
            hot_med = median(CH4_umol_m2_h[hot_moment_iqr == "hot"], na.rm = T),
            quant_95 = quantile(CH4_umol_m2_h, probs = 0.95, na.rm = TRUE),
            min = min(DateTime),
            max = max(DateTime))

annual <- hot_moments %>%
  mutate(Temp_C = sub(" ºC", "", str_extract(as.character(MIU_VALVE), "[0-9]+\\.*[0-9]* ºC")),
         Temp_C = as.numeric(sub("mb", "0", Temp_C)),
         dif = as.numeric(difftime(max, min, "hours"))) %>%
  ggplot(aes(x = Temp_C, 
             y = sum*130/60/1000/1000/dif*365 #convert umol/m2/h to mol/m2
             ))+
  geom_smooth(method = "lm", color = "black")+
  geom_point(aes(fill = MIU_VALVE), color = "black", shape = 21, stroke = 0.5, size = 2)+
  scale_fill_manual(values = color.gradient, name = "Chamber")+
  theme_hm +
  ggpubr::stat_cor(aes(label = ..p.label..),
           label.x.npc = "left", label.y.npc = "top")+
  ylab(expression(atop("Total annual emissions","(mol"~m^-2*")")))+
  xlab("Temperature treatment (ºC above ambient)")+
  theme(legend.position = "bottom",
        legend.title.position = "top")+
  guides(fill = guide_legend(ncol = 2))

top_pct <- hot_moments %>%
  mutate(
    Temp_C = sub(" ºC", "", str_extract(as.character(MIU_VALVE), "[0-9]+\\.*[0-9]* ºC")),
    Temp_C = as.numeric(sub("mb", "0", Temp_C)),
    hm_pct_998 = hot_sum_998/sum * 100,
    hm_pct_99  = hot_sum_99/sum * 100,
    hm_pct_95  = hot_sum_95/sum * 100
  ) %>%
  select(Temp_C, MIU_VALVE, hm_pct_998, hm_pct_99, hm_pct_95) %>%
  pivot_longer(c(hm_pct_95, hm_pct_99, hm_pct_998)) %>%
  mutate(
    name = factor(
      name,
      levels = c("hm_pct_95", "hm_pct_99","hm_pct_998"),
      labels = c("Top 5% of measurements",
                 "Top 1% of measurements",
                 "Top 0.2% of measurements")
    )
  ) %>%
  group_by(name) %>%
  mutate(
    model  = list(lm(value ~ Temp_C)),
    p      = summary(model[[1]])$coefficients[2,4],
    r2     = summary(model[[1]])$r.squared,
    median = median(value)
  ) %>%
  ungroup()

# Create label dataframe (one row per facet)
label_df <- top_pct %>%
  group_by(name) %>%
  summarise(
    p = first(p),
    r2 = first(r2),
    median = first(median),
    label = paste0(
      "  Median = ", round(median, 1), " %",
      "\n  R² = ", round(r2, 2),
      "\n  p = ", round(p, 3)
    )
  )

top_pct_fig <- top_pct %>%
  ggplot(aes(x = Temp_C, y = value)) +
  geom_smooth(method = "lm", color = "black",
              data = ~ subset(.x, p < 0.05)) +
  geom_smooth(method = "lm", color = "grey70",
              data = ~ subset(.x, p >= 0.05),
              lty = "51", alpha = 0.15) +
  geom_point(aes(fill = MIU_VALVE),
             color = "black", shape = 21,
             stroke = 0.5) +
  geom_text(
    data = label_df,
    aes(x = -Inf, y = Inf, label = label),
    hjust = 0, vjust = 1.1,
    inherit.aes = FALSE, size = 2.5
  ) +
  scale_fill_manual(values = color.gradient, name = "Chamber") +
  theme_hm +
  ylab(expression(atop("Percent of annual","emissions"))) +
  xlab("Temperature treatment (ºC above ambient)") +
  guides(fill = guide_legend(ncol = 4)) +
  facet_wrap(~name, scales = "free_y")

top_pct_fig2 <- top_pct %>%
  ggplot(aes(x = Temp_C, y = value)) +
  geom_smooth(method = "lm", color = "black",
              data = ~ subset(.x, p < 0.05)) +
  geom_smooth(method = "lm", color = "grey70",
              data = ~ subset(.x, p >= 0.05),
              lty = "51", alpha = 0.15) +
  geom_point(aes(fill = MIU_VALVE),
             color = "black", shape = 21,
             stroke = 0.5) +
  geom_text(
    data = label_df,
    aes(x = -Inf, y = Inf, label = label),
    hjust = 0, vjust = 1.1,
    inherit.aes = FALSE, size = 2.5
  ) +
  scale_fill_manual(values = color.gradient, name = "Chamber") +
  theme_hm +
  ylab("Percent of annual emissions") +
  xlab("Temperature treatment\n(ºC above ambient)") +
  guides(fill = guide_legend(ncol = 4)) +
  facet_wrap(~name, scales = "free_y", ncol = 1)+
  theme(legend.position = "none")

top_pct_comb_annual <- ggpubr::ggarrange(annual, top_pct_fig2, widths = c(1.3, 1), labels = "auto")

ggsave(top_pct_comb_annual, filename = "../figures/top_pct_comb_annual.jpg", 
       dpi = 300, width = 6, height = 4.5)

ggpubr::ggarrange(annual + theme(legend.position = "right"), 
                  top_pct_fig + theme(legend.position = "none"), 
                  nrow = 2, ncol = 1,
                  labels = "auto")

ggsave(filename = "../figures/top_pct_comb_annual.jpg", 
       dpi = 300, width = 6, height = 5)
```

```{r}
hm_df <- part %>%
  #Re-introduce NAs
  mutate(CH4_umol_m2_h = ifelse(is.na(CH4), NA, CH4_umol_m2_h)) %>%
  filter(!is.na(CH4_umol_m2_h)) 

library(slider)

BEFORE <- lubridate::days(15)
AFTER <- lubridate::days(15)

hm_ident <- hm_df %>%
  mutate(CH4_umol_m2_h = ifelse(CH4_umol_m2_h < 0, 0, CH4_umol_m2_h)) %>%
  arrange(MIU_VALVE, DateTime) %>%
  group_by(MIU_VALVE) %>%
  mutate(
    roll_q75 = slide_index_dbl(
      CH4_umol_m2_h,
      DateTime,
      ~ quantile(.x, probs = 0.75, na.rm = TRUE),
      .before = BEFORE,
      .after  = AFTER,
      .complete = T
    ),
    roll_q95 = slide_index_dbl(
      CH4_umol_m2_h,
      DateTime,
      ~ quantile(.x, probs = 0.95, na.rm = TRUE),
      .before = BEFORE,
      .after  = AFTER,
      .complete = T
    ),
    roll_q99 = slide_index_dbl(
      CH4_umol_m2_h,
      DateTime,
      ~ quantile(.x, probs = 0.95, na.rm = TRUE),
      .before = BEFORE,
      .after  = AFTER,
      .complete = T
    ),
    roll_iqr = slide_index_dbl(
      CH4_umol_m2_h,
      DateTime,
      ~ IQR(.x, na.rm = TRUE),
      .before = BEFORE,
      .after  = AFTER,
      .complete = T
    ),
    hot_moment_quant_95 = ifelse(
      CH4_umol_m2_h > roll_q95,
      "hot", "not"
    ),
    hot_moment_quant_99 = ifelse(
      CH4_umol_m2_h > roll_q99,
      "hot", "not"
    ),
    hot_moment_iqr = ifelse(
      CH4_umol_m2_h > (roll_q75 + 1.5 * roll_iqr),
      "hot", "not"
    ),
    roll_sum = slide_index_dbl(
      CH4_umol_m2_h,
      DateTime,
      ~ sum(.x, na.rm = TRUE),
      .before = BEFORE,
      .after  = AFTER,
      .complete = T
    ),
    hm_quant_99 = ifelse(hot_moment_quant_99 == "hot",
                      CH4,
                      0),
    hm_quant_95 = ifelse(hot_moment_quant_95 == "hot",
                      CH4_umol_m2_h,
                      0),
    hm_iqr = ifelse(hot_moment_iqr == "hot",
                      CH4_umol_m2_h,
                      0),
    roll_sum_iqr = slide_index_dbl(
      hm_iqr, DateTime,
      ~ sum(.x, na.rm = TRUE), 
      .before = BEFORE, .after = AFTER,
      .complete = T
    ),
    roll_sum_quant_99 = slide_index_dbl(
      hm_quant_99, DateTime,
      ~ sum(.x, na.rm = TRUE), 
      .before = BEFORE, .after = AFTER,
      .complete = T
    ),
    roll_sum_quant_95 = slide_index_dbl(
      hm_quant_95, DateTime,
      ~ sum(.x, na.rm = TRUE), 
      .before = BEFORE, .after = AFTER,
      .complete = T
    ),
    roll_prop_iqr = roll_sum_iqr/roll_sum,
    roll_prop_quant_99 = roll_sum_quant_99/roll_sum,
    roll_prop_quant_95 = roll_sum_quant_95/roll_sum,
  ) 

hot_moments <- hm_ident %>%
  group_by(date) %>%
  filter(length(unique(MIU_VALVE[!is.na(hot_moment_iqr)])) == 12) %>%
  group_by(MIU_VALVE) %>%
  summarize(
    median   = median(CH4_umol_m2_h, na.rm = TRUE),
    mean     = mean(CH4_umol_m2_h, na.rm = TRUE),
    sum      = sum(CH4_umol_m2_h, na.rm = TRUE),
    hot_sum  = sum(CH4_umol_m2_h[hot_moment_quant_99 == "hot"], na.rm = TRUE),
    hot_med  = median(CH4_umol_m2_h[hot_moment_quant_99 == "hot"], na.rm = TRUE),
    quant_95 = quantile(CH4_umol_m2_h, probs = 0.95, na.rm = TRUE)
  )

hm <- hot_moments %>%
  ggplot(aes(x = median, y = hot_sum/sum, fill = MIU_VALVE))+
  geom_point(color = "black", shape = 21, stroke = 0.5, size = 2)+
  scale_fill_manual(values = color.gradient, name = "Chamber")+
  egg::theme_article()+
  ylab("Sum of hot moment fluxes (top 5%)")+
  xlab("Median flux")+
  theme(panel.grid.major = element_line(color = "grey93", linewidth = 0.5),
        panel.grid.minor = element_line(color = "grey95", linewidth = 0.25),
        legend.position = "bottom",
        legend.title.position = "top")+
  guides(fill = guide_legend(ncol = 2))

hm

hm_ident %>%
  ggplot(aes(x = DateTime, y = CH4_umol_m2_h))+
  geom_point(size = 0.5, data = . %>% filter(hot_moment_quant_99 == "not"))+
  geom_point(size = 0.5, data = . %>% filter(hot_moment_quant_99 == "hot"), color = "red")+
  facet_wrap(~MIU_VALVE, scales = "free")

hm_ident %>%
  filter(date %in% seq.Date(as.Date("2025-11-01"), as.Date("2025-11-20"), by = "1 day")) %>%
  ggplot(aes(x = DateTime, y = CH4_umol_m2_h))+
  geom_point(size = 0.5, data = . %>% filter(hot_moment_quant_99 == "not"))+
  geom_line()+
  geom_point(size = 0.5, data = . %>% filter(hot_moment_quant_99 == "hot"), color = "red")+
  facet_wrap(~MIU_VALVE, scales = "free")

prop <- hm_ident %>%
  ggplot(aes(x = DateTime, y = roll_sum_quant_95/roll_sum, color = MIU_VALVE))+
  geom_line(size = 0.3)+
  ylab("Hot moment fluxes as proportion of total")+
  #coord_cartesian(ylim = c(0,1))+
  scale_color_manual(values = color.gradient, name = "Chamber")+
  egg::theme_article()+
  #facet_wrap(~MIU_VALVE)+
  theme(legend.position = "bottom")

prop
```


Variability increases with median (Figure 3- option 3)

```{r}
hm_df <- part

daily_summaries <- hm_df %>%
  mutate(time_join = round_date(DateTime, "15 minute")) %>%
  left_join(met, 
            by = c("time_join" = "TIMESTAMP")) %>%
  mutate(CH4_umol_m2_h = ifelse(CH4_umol_m2_h < 0, 0, CH4_umol_m2_h)) %>%
  group_by(MIU_VALVE, date) %>%
  summarize(IQR = IQR(CH4_umol_m2_h, na.rm = T),
            median = median(CH4_umol_m2_h, na.rm = T),
            max = max(CH4_umol_m2_h, na.rm = T),
            min = min(CH4_umol_m2_h, na.rm = T),
            tot = sum(CH4_umol_m2_h, na.rm = T),
            top = max(CH4_umol_m2_h, na.rm = T),
            prop = max/tot,
            q_25 = quantile(CH4_umol_m2_h, probs = 0.25, na.rm = T),
            q_95 = quantile(CH4_umol_m2_h, probs = 0.95, na.rm = T),
            GPP = median(GPP_filled, na.rm = T),
            AirT = median(Ta, na.rm = T))

daily_summaries %>%
  ggplot(aes(x = median, y = IQR, color = AirT))+
  geom_point()+
  facet_wrap(~MIU_VALVE)+
  theme_hm +
  scale_color_viridis_c()

daily_summaries %>%
  ggplot(aes(x = median, y = IQR, color = AirT))+
  geom_smooth(aes(color = MIU_VALVE), method = "lm")+
  theme_hm +
  scale_color_manual(values = color.gradient, name = "Chamber")

daily_summaries %>%
  ggplot(aes(x = median, y = IQR))+
  geom_point()+
  geom_smooth(aes(color = MIU_VALVE), method = "lm")+
  facet_wrap(~MIU_VALVE)+
  theme_hm +
  scale_color_manual(values = color.gradient, name = "Chamber")+
  theme(legend.position = "none")


daily_summaries %>%
  ggplot(aes(x = median, y = IQR))+
  geom_point(aes(color = MIU_VALVE), shape = 21)+
  scale_color_manual(values = color.gradient, name = "Chamber")+
  theme_hm 

daily_summaries %>%
  #filter(median > 10) %>%
  ggplot(aes(x = median, y = prop))+
  geom_point()+
  geom_smooth(aes(color = MIU_VALVE), method = "lm")+
  facet_wrap(~MIU_VALVE)+
  theme_hm +
  scale_color_manual(values = color.gradient, name = "Chamber")+
  theme(legend.position = "none")

daily_summaries %>%
  ggplot(aes(x = q_25, y = q_95))+
  geom_point()+
  geom_smooth(aes(color = MIU_VALVE), method = "lm")+
  facet_wrap(~MIU_VALVE)+
  scale_color_manual(values = color.gradient, name = "Chamber")+
  theme_hm +
  theme(legend.position = "none")

daily_summaries %>%
  mutate(month = month(date),
         adj = as.numeric(sub(" ºC", "", str_extract(MIU_VALVE, "[0-9]+\\.*[0-9]* ºC"))),
         AirT = AirT+adj) %>%
  ggplot(aes(x = AirT, y = median, color = MIU_VALVE))+
  geom_point(shape = 21)+
  geom_smooth(method = "lm")+
  theme_hm +
  scale_color_manual(values = color.gradient, name = "Chamber")+
  facet_wrap(~month, scales = "free")
```

Because of hot moments, manual sampling underestimates fluxes (Figure 4)

```{r}
reps <-  100
samples_per_week <- 1

df <- hm_df %>%
  mutate(week = week(date)) 

calc_mean_flux <- function(ch4_vals, n){
  vals <- sample(ch4_vals, n)
  return(mean(vals))
}

df_out_daytime <- df %>%
  filter(hour(DateTime) >= 9,
         hour(DateTime) <= 15) %>%
  group_by(week) %>%
  cross_join(data.frame(rep = 1:reps)) %>%
  group_by(week, MIU_VALVE, rep) %>%
  summarize(flux = calc_mean_flux(CH4_umol_m2_h, samples_per_week)) %>%
  mutate(time = "daytime")

df_out_fullday <- df %>%
  group_by(week) %>%
  cross_join(data.frame(rep = 1:reps)) %>%
  group_by(week, MIU_VALVE, rep) %>%
  summarize(flux = calc_mean_flux(CH4_umol_m2_h, samples_per_week)) %>% 
  mutate(time = "full day")

df_out_rain <- df %>%
  mutate(time_join = round_date(DateTime, "15 minute")) %>%
  left_join(met %>%
              arrange(TIMESTAMP) %>%
              mutate(
                Rain_cm_roll =
                  slider::slide_dbl(Rain_cm_Tot,
                            sum,
                            .before = 3,
                            .complete = TRUE)
              ), 
            by = c("time_join" = "TIMESTAMP")) %>%
  filter(Rain_cm_roll == 0) %>%
  group_by(week) %>%
  cross_join(data.frame(rep = 1:reps)) %>%
  group_by(week, MIU_VALVE, rep) %>%
  summarize(flux = calc_mean_flux(CH4_umol_m2_h, samples_per_week)) %>% 
  mutate(time = "rain")

df_out_growth <- df %>%
  filter(month(DateTime) %in% 5:9) %>%
  group_by(week) %>%
  cross_join(data.frame(rep = 1:reps)) %>%
  group_by(week, MIU_VALVE, rep) %>%
  summarize(flux = calc_mean_flux(CH4_umol_m2_h, samples_per_week)) %>%
  mutate(time = "growth")

df_out_wl <- df %>%
  mutate(time_join = round_date(DateTime, "15 minute")) %>%
  filter(Depth_cm < 0) %>%
  group_by(week) %>%
  cross_join(data.frame(rep = 1:reps)) %>%
  group_by(week, MIU_VALVE, rep) %>%
  summarize(flux = calc_mean_flux(CH4_umol_m2_h, samples_per_week)) %>%
  mutate(time = "wl")

df_out <- df_out_daytime  %>%
  bind_rows(df_out_fullday) %>%
  #bind_rows(df_out_rain) %>%
  bind_rows(df_out_growth) %>%
  bind_rows(df_out_wl)

to_plot <- df_out %>%
  group_by(MIU_VALVE, rep, time) %>%
  summarize(mean = mean(flux, na.rm = T)) %>%
  left_join(hot_moments %>% 
              rename(tot_real = sum,
                     mean_real = mean)) %>%
  group_by(MIU_VALVE, time) %>%
  summarize(median = median(mean/mean_real * 100 - 100))

sims <- to_plot %>%
  mutate(time = factor(time, 
                       levels = c("full day", "wl", "daytime", "growth"),
                       labels = c("Weekly (all)", "No standing\nwater",
                                  "Daytime", "Growing\nseason"))) %>%
  ggplot(aes(x = time, y = median))+
  geom_hline(yintercept = 0, color = "grey80")+
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(fill = MIU_VALVE),
              width = 0.2, shape = 21, color = "black")+
  scale_fill_manual(values = color.gradient,
                    name = "Chamber")+
  theme_hm +
  theme(axis.title.x = element_blank(),
        legend.title.position = "top")+
  ylab("Percent difference in mean flux\nbetween manual and high-frequency\nsampling (median of 100 simulations)")
sims
ggsave("../figures/Figure 4 - Simulations.jpg", dpi = 300, width = 6, height = 3)
```

Model hm separately
```{r}
mod_df <- hm_ident %>%
  mutate(time_join = round_date(DateTime, "15 minute")) %>%
  left_join(wl, by = c("time_join" = "TIMESTAMP")) %>%
  left_join(met, by = c("time_join" = "TIMESTAMP")) %>%
  group_by(MIU_VALVE) %>%
  mutate(Depth_change = Depth_cm - lag(Depth_cm))

data_not = mod_df %>%
  filter(hot_moment_iqr == "not") %>%
  select(c(CH4_umol_m2_h, Depth_cm, AirTC_Avg, Depth_change)) %>%
  drop_na() %>%
  ungroup() %>%
  filter(MIU_VALVE == "Ch. 12 (+6.0 ºC)") %>%
  select(-MIU_VALVE)

preds_not <- randomForest::randomForest(CH4_umol_m2_h~Depth_cm + AirTC_Avg + Depth_change,
                                        data = data_not,
                                        importance = T)
randomForest::importance(preds_not)
preds_not$rsq[length(preds_not$rsq)]

data_hot = mod_df %>%
  filter(hot_moment_iqr == "hot") %>%
  select(c(CH4_umol_m2_h, Depth_cm, AirTC_Avg, Depth_change)) %>%
  drop_na() %>%
  ungroup() %>%
  filter(MIU_VALVE == "Ch. 12 (+6.0 ºC)") %>%
  select(-MIU_VALVE)

preds_hot <- randomForest::randomForest(CH4_umol_m2_h~Depth_cm + AirTC_Avg + Depth_change,
                                        data = data_hot,
                                        importance = T)
randomForest::importance(preds_hot)
preds_hot$rsq[length(preds_not$rsq)]

data_class = mod_df %>%
  select(c(Depth_cm, AirTC_Avg, Depth_change, hot_moment_iqr)) %>%
  drop_na() %>%
  ungroup() %>%
  filter(MIU_VALVE == "Ch. 12 (+6.0 ºC)") %>%
  select(-MIU_VALVE) %>%
  mutate(hot_moment_iqr = as.factor(hot_moment_iqr))
preds_class <- randomForest::randomForest(hot_moment_iqr~Depth_cm + AirTC_Avg + Depth_change,
                                        data = data_class,
                                        importance = T,
                                        classwt = c(not = 1, hot = 19))
randomForest::importance(preds_class)
preds_class$err.rate[nrow(preds_class$err.rate),]

mod_df %>%
  filter(hot_moment_iqr == "not") %>%
  ggplot(aes(x = DateTime, y = CH4_umol_m2_h, color = MIU_VALVE))+
  geom_point(size = 0.5) +
  geom_line()

patchwork::wrap_plots(mod_df %>%
                        filter(hot_moment_iqr == "not") %>%
                        ggplot(aes(x = DateTime, y = CH4_umol_m2_h, color = MIU_VALVE))+
                        geom_point(size = 0.5) +
                        geom_line()+
                        guides(color = guide_legend(ncol = 4))+
                        theme_hm +
                        theme(axis.text.x = element_blank(),
                              axis.ticks.x = element_blank(),
                              legend.position = "none"), 
                        annual_AT +
                          theme(axis.text.x = element_blank(),
                                axis.ticks.x = element_blank(),
                                legend.position = "none"), 
                        annual_evi +
                          theme(legend.position = "none"),
                      ncol = 1)

mod_df %>%
  filter(!is.na(hot_moment_iqr)) %>%
  ggplot(aes(x = DateTime, y = MIU_VALVE))+
  geom_tile(aes(color = hot_moment_iqr,
                fill = hot_moment_iqr))+
  theme_hm +
  scale_color_manual(values = c("hot" = "red", "not" = "grey95"))+
  scale_fill_manual(values = c("hot" = "red", "not" = "grey95"))
```

```{r}

```

