---
title: "Drivers and CH4"
author: "Abby Lewis"
date: "2024-08-22"
output: html_document
---

Methods from Ari's paper and this one https://agupubs.onlinelibrary.wiley.com/doi/full/10.1002/2015JG003054

Load data and packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(data.table)
library(waveslim)
library(DescTools)
source("../R/wavelet_functions.R")
source("../R/make_lm_label.R")
source("../analysis/hm_style.R")

ch4 <- read_csv(here::here("processed_data","L2- partitioned_and_gap_filled.csv")) %>%
  filter(CH4_filled < 1) %>% #Removing n = 1 flux from chamber 12. Probably good to add an si figure showing what happens if we don't do this
  mutate(CH4_filled = CH4_filled * (60 * 60)) # h instead of s
```

Run wavelet

```{r}
drivers <- c(
  "AirTC_Avg",
  "Depth_cm",
  "GPP_filled",
  "Salinity",
  "Reco"
)

driver_labels <- c(
  AirTC_Avg = "Air temperature",
  Depth_cm = "Water depth",
  GPP_filled = "GPP",
  GPP_unfilled = "GPP",
  Salinity = "Salinity",
  Reco = "Reco"
)

driver_colors <- c(
  AirTC_Avg = "#E76D66",
  Depth_cm = "#2176AB",
  GPP_filled = "#3BAA5D",
  GPP_unfilled = "#3BAA5D",
  Salinity = "#6A4C93",
  Reco = "#264653"
)

#Apply across all treatments and sites
grid <- ch4 %>%
  pivot_longer(cols = c("CH4_filled", drivers),
               #NOTE: I've also tried EVI and PAR but they don't add much
               names_to = "var_name", values_to = "value")

dfs <- map(unique(grid$var_name), 
           double_map, 
           ch4 = ch4, 
           treatment = unique(ch4$MIU_VALVE),
           timestep_s = 610*12) %>%
  bind_rows()

#Re-introduce NAs
nas_to_add <- ch4 %>%
  mutate(na_ch4 = ifelse(is.na(CH4), T, F),
         na_gpp = ifelse(is.na(GPP), T, F)) %>%
  select(TIMESTAMP, MIU_VALVE, na_ch4, na_gpp) %>%
  distinct()

#Group wavelet scales into interpretable bins
discrete_scales <- dfs %>%
  mutate(
    name = as.character(name),
    num_value = as.numeric(gsub("[^0-9]", "", name)),
    grouped_scale = case_when(
      grepl("days", name) & num_value < 30 ~ "multiday",
      grepl("hours", name) & num_value < 10 ~ "hourly",
      grepl("hours", name) & num_value >= 10 ~ "daily",
      TRUE ~ "seasonal"
    )
  ) %>%
  select(-num_value) %>%
  group_by(grouped_scale, MIU_VALVE, TIMESTAMP, var_name) %>%
  summarize(value = sum(value), .groups = "drop") %>%
  pivot_wider(names_from = var_name, values_from = value) %>%
  left_join(nas_to_add) %>%
  mutate(CH4_unfilled = ifelse(na_ch4, NA, CH4_filled),
         GPP_unfilled = ifelse(na_gpp, NA, GPP_filled)) %>%
  select(-na_ch4, -na_gpp)

#Separate explanatory variables
discrete_scales_exp <- discrete_scales %>%
  dplyr::select(-c(CH4_unfilled, CH4_filled, GPP_unfilled)) %>%
  pivot_longer(cols = -c(grouped_scale, MIU_VALVE, TIMESTAMP), 
               names_to = "exp_var", values_to = "exp_var_value") %>%
  filter(!is.na(exp_var_value)) %>%
  rename(grouped_scale_exp = grouped_scale)
```

Plot overall variance

```{r}
grouped_pct_var <- dfs %>%
  mutate(
    name = as.character(name),
    num_value = as.numeric(gsub("[^0-9]", "", name)),
    grouped_scale = case_when(
      grepl("days", name) & num_value < 30 ~ "multiday",
      grepl("hours", name) & num_value < 10 ~ "hourly",
      grepl("hours", name) & num_value >= 10 ~ "daily",
      TRUE ~ "seasonal"
    )
  ) %>%
  select(-num_value) %>%
  group_by(grouped_scale, MIU_VALVE, TIMESTAMP, var_name) %>%
  summarize(value = sum(value), .groups = "drop") %>%
  filter(var_name == "CH4_filled") %>%
  group_by(grouped_scale, MIU_VALVE, var_name) %>%
  summarize(var = var(value, na.rm = T)) %>%
  group_by(MIU_VALVE, var_name) %>%
  mutate(tot_var = sum(var, na.rm = T)) %>%
  ungroup() %>%
  mutate(MIU_VALVE = factor(MIU_VALVE, 
                            levels = 1:12,
                            labels = chamber_levels2),
         grouped_scale = factor(grouped_scale,
                                levels = c("hourly", "daily", "multiday", "seasonal")),
         Temp_C = sub(" ºC", "", str_extract(as.character(MIU_VALVE), "[0-9]+\\.*[0-9]* ºC")),
         Temp_C = as.numeric(sub("mb", "0", Temp_C)))
  
scale_labels <- make_lm_label(
  data = grouped_pct_var %>% mutate(sd = sqrt(var)),
  xvar = "Temp_C",
  yvar = "sd",
  group_var = "grouped_scale"
)

scales <- grouped_pct_var %>%
  mutate(sd = sqrt(var)) %>%
  ggplot(aes(x = Temp_C, y = sd)) +
  geom_smooth(method = "lm", color = "black") +
  geom_point(aes(fill = MIU_VALVE), size = 2, shape = 21) +
  geom_text(
    aes(x = -Inf, y = Inf, label = label),
    hjust = 0, vjust = 1.1,
    inherit.aes = FALSE,
    size = 2.5, 
    data = scale_labels
  ) +
  scale_fill_manual(values = color.gradient, name = "Chamber") +
  facet_wrap(~grouped_scale) +
  theme_hm +
  ylab(expression("Standard deviation (µmol"~m^2~h^-1*")")) +
  xlab("Temperature treatment (ºC above ambient)")

d_to_s_data <- grouped_pct_var %>%
  group_by(MIU_VALVE, Temp_C) %>%
  summarise(
    d_to_s = sqrt(var[grouped_scale == "daily"]) /
             sqrt(var[grouped_scale == "seasonal"]),
    .groups = "drop"
  )

d_to_s_label <- make_lm_label(
  data = d_to_s_data,
  xvar = "Temp_C",
  yvar = "d_to_s"
)

d_to_s <- d_to_s_data %>%
  ggplot(aes(x = Temp_C, y = d_to_s)) +
  geom_smooth(method = "lm", color = "black") +
  geom_point(aes(fill = MIU_VALVE), size = 2, shape = 21) +
  geom_text(
    aes(x = -Inf, y = Inf, label = label),
    hjust = 0, vjust = 1.1,
    inherit.aes = FALSE,
    size = 2.5, 
    data = d_to_s_label
  ) +
  scale_fill_manual(values = color.gradient, name = "Chamber") +
  theme_hm +
  ylab("Daily : seasonal") +
  xlab("Temperature treatment\n(ºC above ambient)")

comb <- ggpubr::ggarrange(scales + theme(legend.position = "none"),
          d_to_s + theme(legend.position = "bottom") +
            guides(fill = guide_legend(ncol = 2,
                                       title.position = "top",
                                       theme = theme(legend.text = element_text(size = 8),
                                                     legend.key.spacing.y = unit(0, "cm"),
                                                     legend.key.spacing.x = unit(0, "cm"))))+
            ylab("Daily : seasonal")+
            ggtitle(""), 
          nrow = 1, widths = c(1.6,1), labels = "auto")

comb

ggsave(comb, filename = "../figures/variability_combined.png", width = 7, height = 4)
```

Figure showing altered variability (maybe for SI?)

```{r}
scales <- discrete_scales %>%
  filter(grouped_scale == "seasonal" | 
           month(TIMESTAMP) == 7 & day(TIMESTAMP) %in% 10:17 & grouped_scale == "multiday" |
           month(TIMESTAMP) == 7 & day(TIMESTAMP) %in% 10:17 & grouped_scale == "daily" |
           month(TIMESTAMP) == 7 & day(TIMESTAMP) %in% 10:17 & grouped_scale == "hourly"
           ) %>%
  mutate(MIU_VALVE = factor(MIU_VALVE,
                            levels = 1:12,
                            labels = chamber_levels2),
         grouped_scale = factor(grouped_scale,
                                levels = rev(c("hourly", "daily", "multiday", "seasonal")))) %>%
  ggplot(aes(x = TIMESTAMP, y = CH4_filled, color = MIU_VALVE))+
  geom_line()+
  ylab(expression("Methane flux (µmol"~m^2~h^-1*")"))+
  facet_wrap(~grouped_scale, scales = "free_x")+
  scale_color_manual(values = color.gradient, name = "Chamber")+
  theme_hm+
  scale_x_datetime(
  labels = function(x) {
    ifelse(
      format(x, "%d") == "01",
      format(x, "%b"),
      format(x, "%b %d")
    )
  }
)+
  theme(axis.title.x = element_blank())

scales

ggsave(scales, filename = "../figures/wavelet_example.png", width = 5, height = 3.5)
```

Spearman

```{r}
# Calculate Spearman correlation
spearman_corrs <- discrete_scales %>%
  dplyr::select(c(grouped_scale, MIU_VALVE, TIMESTAMP, CH4_filled)) %>%
  left_join(discrete_scales_exp, relationship = "many-to-many") %>%
  filter(!is.na(exp_var_value),
         !is.na(CH4_filled)) %>%
  group_by(MIU_VALVE, grouped_scale, exp_var, grouped_scale_exp) %>%
  summarize(
    spearman_rho = cor(exp_var_value, CH4_filled, 
                       method = "spearman", 
                       use = "complete.obs"),
    spearman_rho2 = spearman_rho^2,
    .groups = "drop"
  ) %>%
  mutate(grouped_scale = factor(grouped_scale,
                                levels = c("hourly", "daily", "multiday", "seasonal")))

relative_corr <- spearman_corrs %>%
  filter(grouped_scale == grouped_scale_exp,
         !exp_var == "evi_predicted") %>%
  mutate(
    abs_rho = abs(spearman_rho),
    MIU_VALVE = factor(MIU_VALVE,
                       levels = 1:12,
                       labels = chamber_levels2)
  )

#Maybe for SI
ggplot(relative_corr,
       aes(x = grouped_scale,
           y = abs_rho,
           color = exp_var,
           group = exp_var)) +
  geom_point(size = 0.5) +
  geom_line() +
  facet_wrap(~MIU_VALVE) +
  xlab("Wavelet scale of variation") +
  ylab("Absolute Spearman correlation (|ρ|)") +
  scale_color_manual(values = driver_colors[drivers],
                     labels = driver_labels[drivers],
                     name = "Explanatory variable") +
  theme_hm +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

relative_corr %>%
  ggplot(aes(x = grouped_scale,
             y = abs_rho,
             color = exp_var,
             group = exp_var)) +
  geom_line(alpha = 0.1, aes(group = paste0(MIU_VALVE, exp_var)))+
  geom_line(data = . %>% group_by(exp_var, grouped_scale) %>%
              summarize(abs_rho = mean(abs_rho)),
            linewidth = 1.5)+
  xlab("Wavelet scale of variation") +
  ylab("Spearman correlation |ρ|)") +
  scale_color_manual(values = driver_colors[drivers],
                     labels = driver_labels[drivers],
                     name = "Explanatory variable") +
  theme_hm + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

spear <- relative_corr %>%
  ggplot(aes(x = grouped_scale,
             y = spearman_rho,
             color = exp_var,
             group = exp_var)) +
  geom_hline(color = "grey50", yintercept = 0)+
  geom_line(alpha = 0.2, aes(group = paste0(MIU_VALVE, exp_var)))+
  geom_line(data = . %>% group_by(exp_var, grouped_scale) %>%
              summarize(spearman_rho = mean(spearman_rho)),
            linewidth = 1.5)+
  scale_x_discrete(expand = c(0, 0.4))+
  xlab("Wavelet scale of variation") +
  ylab("Spearman correlation (ρ)") +
  scale_color_manual(values = driver_colors[drivers],
                     labels = driver_labels[drivers],
                     name = "Explanatory variable") +
  guides(color = guide_legend(ncol = 2))+
  theme_hm +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

spear

ggsave(spear, filename = "../figures/spearman_correlations.png", width = 3, height = 4)
```

Correlation among chambers

```{r}
to_plot <- discrete_scales %>%
  ungroup() %>%
  select(TIMESTAMP, grouped_scale, MIU_VALVE, CH4_filled) %>%
  mutate(TIMESTAMP = round_date(TIMESTAMP, "2 hours")) %>%
  pivot_wider(names_from = MIU_VALVE, values_from = CH4_filled) %>%
  group_by(grouped_scale) %>%
  summarise(
    cor_matrix = list(
      cor(across(where(is.numeric)),
          use = "complete.obs",
          method = "spearman")
    )
  ) %>%
  mutate(cor_df = map(cor_matrix, ~ as.data.frame(as.table(.x)))) %>%
  unnest(cor_df)%>%
  mutate(grouped_scale = factor(grouped_scale,
                                levels = c("hourly", "daily", "multiday", "seasonal")))

dfs %>%
  filter(name %in% c("16 hours"),
         var_name == "CH4_filled",
         month(TIMESTAMP) == 9) %>%
  group_by(MIU_VALVE, TIMESTAMP) %>%
  summarize(value = sum(value)) %>%
  ggplot(aes(x = TIMESTAMP, y = value)) +
  geom_line()+
  facet_wrap(~MIU_VALVE)

unique(dfs$name)

geom <- read_csv(here::here("Raw_data", "genx_lat_lon.csv"))
library(geosphere)

# Create all combinations
pairs <- expand.grid(
  Var1 = geom$chamber,
  Var2 = geom$chamber,
  stringsAsFactors = FALSE
)

# Join coordinates and compute distance
result <- pairs %>%
  left_join(geom, by = c("Var1" = "chamber")) %>%
  rename(lat1 = latitude, lon1 = longitude) %>%
  left_join(geom, by = c("Var2" = "chamber")) %>%
  rename(lat2 = latitude, lon2 = longitude) %>%
  rowwise() %>%
  mutate(Distance = distHaversine(
    c(lon1, lat1),
    c(lon2, lat2)
  )) %>%
  ungroup() %>%
  select(Var1, Var2, Distance) %>%
  mutate(Var1 = factor(as.numeric(sub("ch", "", Var1))),
         Var2 = factor(as.numeric(sub("ch", "", Var2))))

to_plot %>%
  left_join(result) %>%
  filter(!Distance == 0) %>%
  group_by(grouped_scale, Var1) %>%
  mutate(p = summary(lm(Freq~Distance))$coefficients[2,4]) %>%
  ggplot(aes(x = Distance, y = Freq, color = Var1))+
  geom_point(shape = 21, alpha = 0.5)+
  geom_smooth(se = F, method = "lm", data = . %>% filter(p < 0.05))+
  geom_line(stat = "smooth", se = F, method = "lm", 
            data = . %>% filter(p >= 0.05), alpha = 0.25, size = 1, lty = "51")+
  facet_grid(grouped_scale~Var1, scales = "free_y")+
  theme_hm +
  scale_color_manual(values = color.gradient)

to_plot %>%
  ggplot(aes(x = grouped_scale, y = Freq, color = Var2, group = Var2)) +
  geom_line() +
  facet_wrap(~Var1)+
  theme_hm +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_color_manual(name = "Chamber", 
                     values = color.gradient) 

to_plot %>%
  filter(!Var1 == Var2) %>%
  ggplot(aes(x = Var1, y = Freq, color = grouped_scale, group = grouped_scale)) +
  geom_line() +
  facet_wrap(~Var2)+
  theme_hm +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Correlation among chambers first dif

```{r}
to_plot_dif <- discrete_scales %>%
  ungroup() %>%
  select(TIMESTAMP, grouped_scale, MIU_VALVE, CH4_filled) %>%
  mutate(TIMESTAMP = round_date(TIMESTAMP, "2 hours")) %>%
  group_by(MIU_VALVE, grouped_scale) %>%
  arrange(TIMESTAMP) %>%
  mutate(CH4_filled = CH4_filled - lag(CH4_filled)) %>%
  ungroup() %>%
  pivot_wider(names_from = MIU_VALVE, values_from = CH4_filled) %>%
  group_by(grouped_scale) %>%
  summarise(
    cor_matrix = list(
      cor(across(where(is.numeric)),
          use = "complete.obs",
          method = "spearman")
    )
  ) %>%
  mutate(cor_df = map(cor_matrix, ~ as.data.frame(as.table(.x)))) %>%
  unnest(cor_df) %>%
  mutate(grouped_scale = factor(grouped_scale,
                                levels = c("hourly", "daily", "multiday", "seasonal")),
         Var1 = factor(Var1, levels = 1:12),
         Var2 = factor(Var2, levels = 1:12))

to_plot_dif %>%
  ggplot(aes(x = grouped_scale, y = Freq, color = Var2, group = Var2)) +
  geom_line() +
  facet_wrap(~Var1)+
  theme_hm +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_color_manual(name = "Chamber", 
                     values = color.gradient) 

to_plot_dif %>%
  left_join(result) %>%
  filter(!Distance == 0) %>%
  group_by(grouped_scale, Var1) %>%
  mutate(p = summary(lm(Freq~Distance))$coefficients[2,4]) %>%
  ggplot(aes(x = Distance, y = Freq, color = Var1))+
  geom_point(shape = 21, alpha = 0.5)+
  geom_smooth(se = F, method = "lm", data = . %>% filter(p < 0.05))+
  geom_line(stat = "smooth", se = F, method = "lm", 
            data = . %>% filter(p >= 0.05), alpha = 0.25, size = 1, lty = "51")+
  facet_grid(grouped_scale~Var1, scales = "free_y")+
  theme_hm +
  scale_color_manual(values = color.gradient)

to_plot_dif %>%
  left_join(result) %>%
  filter(!Distance == 0) %>%
  group_by(grouped_scale, Var1) %>%
  mutate(p = summary(lm(Freq~Distance))$coefficients[2,4]) %>%
  ggplot(aes(x = Distance, y = Freq, color = Var1))+
  geom_point(shape = 21, alpha = 0.5)+
  geom_smooth(se = F, method = "lm", data = . %>% filter(p < 0.05))+
  geom_line(stat = "smooth", se = F, method = "lm", 
            data = . %>% filter(p >= 0.05), alpha = 0.25, size = 1, lty = "51")+
  facet_grid(grouped_scale~Var1, scales = "free_y")+
  theme_hm +
  scale_color_manual(values = color.gradient)

temps <- c(0,0,0.75, 1.5, 2.25, 2.25, 3.0, 3.75, 3.75, 4.5, 5.25, 6.0)

to_plot_dif %>%
  filter(!Var1 == Var2) %>%
  mutate(Temp1 = temps[as.numeric(Var1)],
         Temp2 = temps[as.numeric(Var2)],
         Distance = abs(Temp1-Temp2)) %>%
  group_by(grouped_scale, Var1) %>%
  mutate(p = summary(lm(Freq~Distance))$coefficients[2,4]) %>%
  ggplot(aes(x = Distance, y = Freq, color = Var1))+
  geom_point(shape = 21, alpha = 0.5)+
  geom_smooth(se = F, method = "lm", data = . %>% filter(p < 0.05))+
  geom_line(stat = "smooth", se = F, method = "lm", 
            data = . %>% filter(p >= 0.05), alpha = 0.25, size = 1, lty = "51")+
  facet_grid(grouped_scale~Var1, scales = "free_y")+
  theme_hm +
  scale_color_manual(values = color.gradient)

to_plot_dif %>%
  filter(!Var1 == Var2) %>%
  mutate(Temp1 = temps[as.numeric(Var1)],
         Temp2 = temps[as.numeric(Var2)],
         Distance = Temp2-Temp1) %>%
  group_by(grouped_scale, Var1) %>%
  mutate(p = summary(lm(Freq~Distance))$coefficients[2,4]) %>%
  ggplot(aes(x = Distance, y = Freq))+
  geom_point(aes(color = Var1), shape = 21, alpha = 0.5)+
  #geom_smooth(aes(color = Var1), se = F, method = "lm", data = . %>% filter(p < 0.05))+
  geom_smooth(aes(color = Var1), se = F, method = "gam",
              formula = y ~s(x, bs = "cs", fx = T, k = 3))+
  geom_smooth(color = "black", se = F, method = "gam")+
  #geom_line(aes(color = Var1), stat = "smooth", se = F, method = "lm", 
  #          data = . %>% filter(p >= 0.05), alpha = 0.25, size = 1, lty = "51")+
  facet_wrap(~grouped_scale, scales = "free_y")+
  theme_hm +
  scale_color_manual(values = color.gradient, name = "Chamber")+
  xlab("Temperature difference (ºC)")+
  ylab("Correlation between chambers")


to_plot_dif %>%
  #filter(grouped_scale %in% c("daily","seasonal")) %>%
  filter(!Var1 == Var2) %>%
  ggplot(aes(x = Var1, y = Freq, color = Var2, group = Var2)) +
  geom_line() +
  facet_wrap(~grouped_scale)+
  theme_hm +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_color_manual(values = color.gradient)
```

Random forest
- Benefit of this over MI is that MI requires binning the data, which loses information especially for chamber 12
- Spearmans correlation is simpler
- This takes a while to run

```{r}
library(caret)
library(ranger)
set.seed(47)

rf_results <- data.frame()
rf_importance <- data.frame()

scales   <- unique(discrete_scales$grouped_scale)
chambers <- unique(discrete_scales$MIU_VALVE)

# 5-fold CV setup
ctrl <- trainControl(
  method = "cv",
  number = 5,
  savePredictions = "final"
)

for (chamber in chambers) {
  
  cat("Processing MIU_VALVE:", chamber, "\n")
  
  df_valve <- discrete_scales %>%
    filter(MIU_VALVE == chamber) %>%
    drop_na(CH4_filled)
  
  for (scale_name in scales) {
    
    df_scale <- df_valve %>%
      filter(grouped_scale == scale_name) %>%
      as.data.frame()
    
    if (nrow(df_scale) < 5) {
      cat("  Skipping scale", scale_name, "- not enough data\n")
      next
    }
    
    # Train RF with caret
    rf_fit <- train(
      x = df_scale[, drivers],
      y = df_scale$CH4_filled,
      method = "ranger",
      trControl = ctrl,
      importance = "permutation",
      num.trees = 500,
      tuneGrid = data.frame(
        mtry = floor(sqrt(length(drivers))),
        splitrule = "variance",
        min.node.size = 5
      )
    )
    
    # Extract CV performance
    res <- rf_fit$resample
    
    rf_results <- rbind(
      rf_results,
      data.frame(
        MIU_VALVE = chamber,
        scale = scale_name,
        mean_rsq  = mean(res$Rsquared),
        sd_rsq    = sd(res$Rsquared),
        mean_rmse = mean(res$RMSE),
        sd_rmse   = sd(res$RMSE)
      )
    )
    
    # Variable importance (from final model)
    imp <- varImp(rf_fit)$importance
    
    rf_importance <- rbind(
      rf_importance,
      data.frame(
        MIU_VALVE = chamber,
        scale = scale_name,
        variable = rownames(imp),
        importance = imp$Overall
      )
    )
    
  }
  
} # end MIU_VALVE loop

# Plot R2 by scale for all chambers
#Note I don't think this means anything
rf_results %>%
  mutate(MIU_VALVE = factor(MIU_VALVE,
                            levels = 1:12,
                            labels = chamber_levels2)) %>%
  ggplot(aes(x = MIU_VALVE, y = mean_rsq, color = MIU_VALVE)) +
  geom_point() +
  geom_line() +
  geom_errorbar(aes(ymin = mean_rsq - sd_rsq, ymax = mean_rsq + sd_rsq), width = 0.2) +
  theme_hm +
  ylab("Cross-validated R²") +
  xlab("Timescale") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_color_manual(name = "Site", values = color.gradient)+
  facet_wrap(~scale, scales = "free")

# Variable importance averaged over folds
rf_importance_summary <- rf_importance %>%
  group_by(MIU_VALVE, scale, variable) %>%
  summarize(median_importance = median(importance), .groups = "drop")

rf_importance_summary %>%
  mutate(MIU_VALVE = factor(MIU_VALVE,
                            levels = 1:12,
                            labels = chamber_levels2),
         scale = factor(scale,
                        levels = unique(discrete_scales$grouped_scale))) %>%
  ggplot(aes(x = variable, y = median_importance, color = MIU_VALVE)) +
  geom_point()+
  facet_wrap(~scale) +
  coord_flip() +
  theme_hm +
  ylab("Permutation importance") +
  xlab("Predictor variable")+
  scale_color_manual(name = "Site", values = color.gradient)

rf_importance_summary %>%
  mutate(MIU_VALVE = factor(MIU_VALVE,
                            levels = 1:12,
                            labels = chamber_levels2),
         scale = factor(scale,
                        levels = levels(discrete_scales$grouped_scale))) %>%
  ggplot(aes(x = scale, y = median_importance, color = MIU_VALVE, group = MIU_VALVE)) +
  geom_line()+
  facet_wrap(~variable) +
  theme_hm +
  ylab("Permutation importance") +
  xlab("Scale")+
  scale_color_manual(name = "Site", values = color.gradient)

```

With lags

```{r}
discrete_scales_exp <- discrete_scales_exp %>%
  group_by(grouped_scale_exp, MIU_VALVE, exp_var) %>%
  mutate(exp_bin = bin_ranked(exp_var_value, N_BINS))

discrete_scales_ch4 <- discrete_scales %>%
  group_by(grouped_scale, MIU_VALVE) %>%
  mutate(ch4_bin = bin_ranked(CH4_filled, N_BINS))

lagged_mut_infs <- map_dfr(unique(discrete_scales$grouped_scale), function(scale_name) {
  
  message("Processing scale: ", scale_name)
  n <- which(levels(discrete_scales$grouped_scale) == scale_name)
  mult <- n*6
  
  lags <- round(seq(-mult, mult, length.out = 9))
  
  # subset this scale only
  exp_data_scale <- discrete_scales_exp %>%
    filter(grouped_scale_exp == scale_name)
  
  ch4_scale <- discrete_scales_ch4 %>%
    filter(grouped_scale == scale_name)
  
  map_dfr(lags, function(lg) {
    
    lagged_data <- exp_data_scale %>%
      group_by(MIU_VALVE, exp_var) %>%
      arrange(TIMESTAMP) %>%
      mutate(
        exp_bin_lag =
          if (lg < 0) dplyr::lag(exp_bin, abs(lg))
          else if (lg > 0) dplyr::lead(exp_bin, lg)
          else exp_bin,
        lag = lg
      ) %>%
      ungroup()
    
    ch4_scale %>%
      left_join(lagged_data,
                by = c("MIU_VALVE", "TIMESTAMP")) %>%
      filter(!is.na(exp_bin_lag),
             !is.na(ch4_bin)) %>%
      group_by(MIU_VALVE, exp_var, lag) %>%
      summarize(
        entropy_y = entropy(ch4_bin),
        mut_inf = MutInf(exp_bin_lag, ch4_bin) / entropy_y,
        grouped_scale = scale_name,
        .groups = "drop"
      )
  })
})

best_lagged_mut_infs <- lagged_mut_infs %>%
  group_by(MIU_VALVE, grouped_scale, exp_var) %>%
  filter(mut_inf == max(mut_inf, na.rm = TRUE)) %>%
  ungroup()

lagged_mut_infs %>%
  mutate(MIU_VALVE = factor(MIU_VALVE,
                            levels = 1:12,
                            labels = chamber_levels2),
         lag=as.numeric(lag)) %>%
  ggplot(aes(x = lag, y = mut_inf, color = MIU_VALVE, group = MIU_VALVE))+
  geom_line()+
  ggh4x::facet_grid2(grouped_scale~exp_var, scales = "free", independent = "x")+
  scale_color_manual(name = "Chamber", values = color.gradient)
```

Ari:
Quantile regressions were fitted for the 0.1, 0.5, and 0.9 quantiles of CH4 fluxes using the nlrq() function within the R package ‘quantreg’ (Koenker, 2023) due to the observed non-linear relationships between CH4 fluxes and the tested predictor variables. The slopes of the fitted conditional quantile regressions were used to estimate the predictor level required to decrease CH4 fluxes by half, based on an exponential decay relationship (i.e., X1/2 = ln(2)/slope). Subsequently, threshold values were estimated as 7 times X1/2, representing a 99% reduction of CH4 fluxes through interactions with increasing predictor levels. We calculated these thresholds for the 50th and 90th percentiles of the conditional distribution of CH4 fluxes, representing the predictor thresholds below which the 50% and 90% of the highest CH4 fluxes occur, respectively.

Partial associations:
https://agupubs.onlinelibrary.wiley.com/doi/full/10.1002/2013WR013845
