---
title: "CH4 analyses"
author: "Abby Lewis"
date: "2026-01-20"
output: html_document
---

## Overview

This file contains code to analyze and visualize the effects of soil warming on methane emissions, including changes in the proportional contribution of extreme fluxes (i.e., top 5%, 1%, and 0.2%). Code chunks should be run in order.

### Table of contents: 

-   Step 1: Load data and packages
-   Step 2: Calculate summary statistics
-   Step 3: Plot data summary (Figure 2)
-   Step 4: Analyze the contribution of extreme fluxes (Figure 3)
-   Step 5: Simulate low-frequency sampling to assess bias (Figure 4)
-   Step 6: Daytime and WL sampling biases (SI)
-   Step 7: Analyze residuals for annual and extreme fluxes


## Step 1: Load data and packages

```{r setup}
# Load packages and functions
library(tidyverse)
source("hm_style.R")
source("make_lm_label.R")
source("calc_bias.R")

# Load data
ch4 <- read_csv(here::here("processed_data", "partitioned_co2.csv")) %>%
  mutate(DateTime_EST = with_tz(DateTime, tz = "EST"),
         Chamber = factor(MIU_VALVE, levels = 1:12, labels = chamber_labels),
         date = as_date(DateTime_EST)
  )

soil_temp <- read_csv(here::here("processed_data", "for_HM_MS","Soil_temp.csv")) %>%
  mutate(DateTime_EST = with_tz(DateTime_EST, tz = "EST"),
         Chamber = factor(Chamber, levels = chamber_labels))
```


```{r}
library(ggh4x)
df <- ch4 %>%
  mutate(
    hour = round(hour(DateTime_EST)/2)*2,
    year = year(DateTime_EST),
    month = month(DateTime_EST)
  ) %>%
  group_by(hour, year, Chamber, month) %>%
  summarize(eb = sum(Ebullition_yn, na.rm = T)/sum(!is.na(Ebullition_yn))*100,
            pressure = mean(Barometric_Pressure_PB110B_Avg, na.rm = T),
            .groups = "drop") %>%
  pivot_longer(c(eb, pressure))

df %>%
  mutate(name = factor(name,
                       levels = c("eb", "pressure"),
                       labels = c("Ebulition (%)", "Mean pressure"))) %>%
  filter(Chamber == "Ch. 12 (+5.7 ºC)") %>%
  ggplot(aes(x = hour, y = value, color = as.factor(month)))+
  geom_point()+
  geom_smooth(se = F)+
  ylab("Proportion of fluxes that are ebullition")+
  facet_grid(name~year, scales = "free")+
  theme_hm

df2 <- ch4 %>%
  mutate(
    hour = round(hour(DateTime_EST)/2)*2,
    year = year(DateTime_EST)
  ) %>%
  filter(month(DateTime_EST) %in% 3:9) %>%
  group_by(hour, year, Chamber) %>%
  summarize(eb = sum(Ebullition_yn, na.rm = T)/sum(!is.na(Ebullition_yn))*100,
            pressure = mean(Barometric_Pressure_PB110B_Avg, na.rm = T),
            .groups = "drop")

df2 %>%
  ggplot(aes(x = hour, y = eb, color = as.factor(year)))+
  geom_point()+
  geom_smooth()+
  ylab("Proportion of fluxes that are ebullition")+
  facet_wrap(~Chamber)+
  theme_hm

ch4 %>%
  filter(Chamber == "Ch. 12 (+5.7 ºC)") %>%
  mutate(doy = yday(DateTime_EST),
         year = as.factor(year(DateTime_EST))) %>%
  ggplot(aes(x = doy, y = Barometric_Pressure_PB110B_Avg, color = year)) +
  geom_smooth()+
  theme_hm
```

