---
title: "CH4 analyses"
author: "Abby Lewis"
date: "2026-01-20"
output: html_document
---

## Overview

This file contains code to analyze and visualize the effects of soil warming on methane emissions, including changes in the proportional contribution of extreme fluxes (i.e., top 5%, 1%, and 0.2%). Code chunks should be run in order.

### Table of contents: 

-   Step 1: Load data and packages
-   Step 2: Calculate summary statistics
-   Step 3: Plot data summary (Figure 2)
-   Step 4: Analyze the contribution of extreme fluxes (Figure 3)
-   Step 5: Simulate low-frequency sampling to assess bias (Figure 4)
-   Step 6: Daytime and WL sampling biases (SI)
-   Step 7: Analyze residuals for annual and extreme fluxes


## Step 1: Load data and packages

```{r setup}
# Load packages and functions
library(tidyverse)
source("hm_style.R")
source("make_lm_label.R")
source("calc_bias.R")

# Load data
ch4 <- read_csv(here::here("01_Data", "GENX_flux_data_interpolated_with_drivers.csv")) %>%
  mutate(DateTime_EST = with_tz(DateTime_EST, tz = "EST"),
         Chamber = factor(Chamber, levels = chamber_labels),
         date = as_date(DateTime_EST)
  )

soil_temp <- read_csv(here::here("01_Data", "Soil_temp.csv")) %>%
  mutate(DateTime_EST = with_tz(DateTime_EST, tz = "EST"),
         Chamber = factor(Chamber, levels = chamber_labels))
```


```{r}
library(ggh4x)
df <- ch4 %>%
  mutate(
    day = ifelse(GPP_umol_m2_h > 0, "day", "night")
  ) %>%
  group_by(date) %>%
  mutate(temp = mean(AirTemp_C, na.rm = T),
         depth = mean(Depth_cm, na.rm = T),
         reco = mean(Reco_umol_m2_h, na.rm = T),
         gpp = mean(GPP_umol_m2_h, na.rm = T)) %>%
  group_by(date, day, Chamber, temp, depth, reco, gpp) %>%
  summarize(CH4 = mean(CH4_umol_m2_h),
            temp = unique(temp),
            depth = unique(depth),
            reco = unique(reco),
            gpp = unique(gpp),
            .groups = "drop") %>%
  pivot_wider(names_from = day, values_from = CH4) %>%
  mutate(yday = yday(date)) %>%
  group_by(Chamber) %>%
  mutate(
    lims = list(range(c(day, night), na.rm = TRUE))
  ) %>%
  ungroup()

lims_list <- df %>%
  arrange(Chamber) %>%
  distinct(Chamber, lims) %>%
  pull(lims)

df_smoothed <- df %>%
  filter(!is.na(day),
         !is.na(night)) %>%
  group_by(Chamber) %>%
  mutate(
    day_smooth = predict(loess(day ~ yday, span = 0.3)),
    night_smooth = predict(loess(night ~ yday, span = 0.3))
  )

ggplot(df, aes(x = night, y = day)) +
  geom_abline(slope = 1) +
  geom_point(aes(color = yday), shape = 21, alpha = 0.2) +
  geom_point(data = df_smoothed, 
             aes(y = day_smooth, x = night_smooth),
             color = "black") +
  geom_point(data = df_smoothed, 
             aes(y = day_smooth, x = night_smooth, color = yday),
             size = 0.5) +
  facet_wrap(~ Chamber, scales = "free") +
  ggh4x::facetted_pos_scales(
    x = lapply(lims_list, \(z)
      scale_x_continuous(limits = z)
    ),
    y = lapply(lims_list, \(z)
      scale_y_continuous(limits = z)
    )
  ) +
  egg::theme_article() +
  theme(aspect.ratio = 1) +
  scale_color_viridis_c()

ggplot(df, aes(x = night, y = day)) +
  geom_abline(slope = 1) +
  geom_point(aes(color = yday), shape = 21, alpha = 0.2) +
  geom_point(data = df_smoothed, 
             aes(y = day_smooth, x = night_smooth),
             color = "black") +
  geom_point(data = df_smoothed, 
             aes(y = day_smooth, x = night_smooth, color = yday),
             size = 0.5) +
  facet_wrap(~ Chamber) +
  egg::theme_article() +
  theme(aspect.ratio = 1) +
  scale_color_viridis_c()

df %>%
  filter(Chamber %in% c("Ch. 1 (+0 ºC)",
                        "Ch. 2 (-0.1 ºC)")) %>%
  ggplot(aes(x = night, y = day)) +
  geom_abline(slope = 1) +
  geom_point(aes(color = yday), shape = 21, alpha = 0.2) +
  geom_point(data = df_smoothed %>%
  filter(Chamber %in% c("Ch. 1 (+0 ºC)",
                        "Ch. 2 (-0.1 ºC)")), 
             aes(y = day_smooth, x = night_smooth),
             color = "black") +
  geom_point(data = df_smoothed %>%
  filter(Chamber %in% c("Ch. 1 (+0 ºC)",
                        "Ch. 2 (-0.1 ºC)")), 
             aes(y = day_smooth, x = night_smooth, color = yday),
             size = 0.5) +
  facet_wrap(~ Chamber) +
  egg::theme_article() +
  theme(aspect.ratio = 1) +
  scale_color_viridis_c()
```

